Understanding Traffic Patterns and Median Speeds: Insights from Toronto’s Speed Display Program

### 2024-12-01 paper/paper.qmd
added model comparison

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

@@

modelsummary::modelsummary(
  list(
    "Bayesian model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

>>>>

### Consideration of alternative models

To compare the prediction accuracy of the Bayesian and Frequentist regression models, the dataset was divided into a training set (80% of the data) and a testing set (20% of the data). Both models were trained on the training set using the same predictors: traffic direction and the count of vehicles exceeding 100 km/h. The Bayesian model incorporated prior beliefs about the parameter estimates, while the Frequentist model relied solely on the observed data for parameter estimation. After training, predictions were generated for the testing set. For the Bayesian model, predictions were derived by averaging the posterior predictive distributions for each observation, while for the Frequentist model, point predictions were directly calculated from the fitted parameters. Prediction accuracy was assessed using two metrics: Root Mean Squared Error (RMSE), which quantifies the average squared difference between predicted and observed values, and Mean Absolute Error (MAE), which measures the average absolute difference.

The results indicate that both models performed almost identically in predicting median speeds. The Bayesian model achieved an RMSE of 9.603 and an MAE of 7.526, while the Frequentist model had an RMSE of 9.605 and an MAE of 7.528. These results suggest that both approaches captured the key patterns in the data with similar levels of accuracy. The slightly lower RMSE and MAE for the Bayesian model indicate a marginal advantage in prediction, potentially due to its ability to incorporate prior uncertainty into the modeling process. However, the differences between the two models are negligible, implying that either method could be effectively used for this analysis.

The Bayesian model is preferred in this analysis for several reasons, despite the minimal differences in prediction accuracy. First, the Bayesian approach allows for the incorporation of prior information, which can help guide the estimation process, particularly in cases where data may be sparse or noisy. This can result in more robust and reliable parameter estimates, as the Bayesian framework explicitly accounts for uncertainty in both the data and the model parameters. Second, the Bayesian model provides a full posterior distribution for each parameter, enabling a richer interpretation of results and a clearer understanding of the uncertainty surrounding predictions. This is particularly advantageous for decision-making in contexts such as traffic safety, where understanding variability and risk is crucial. Lastly, the Bayesian model’s slightly lower RMSE and MAE, while small, suggest a better fit to the data, likely due to its ability to balance the influence of extreme values and outliers. For these reasons, the Bayesian model is selected as the final model for this analysis, as it offers not only comparable prediction accuracy but also a more nuanced and interpretable framework for addressing the research questions.

@@

Bayesian_model <-
  readRDS(file = here::here("models/Bayesian_model.rds"))
```

@@

modelsummary::modelsummary(
  list(
    "Bayesian model" = Bayesian_model
  ),
  statistic = "mad",
  fmt = 2
)
```
